---
title: "Insertion Analysis"
output: 
  html_notebook: default
date: "2023.07.28"
author: "Jennifer Stiens"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
i_am("Tn_seq_project.Rproj")

library(here)
library(tidyverse)

```


Jupyter notebook for all transit work on thoth
/d/in16/u/sj003/men_tnseq/


## Insertion file Analysis

29 July 2023

Analyse insertions and distributions of samples after creation of insertion files (.wig)


```{r summary_stats}
library(tidyverse)
summary <- read_csv("output/sample_stats.csv", col_names=T, show_col_types = F)
summary$menadione <- as.factor(summary$menadione)
summary$experiment <- as.factor(summary$experiment)
#plot number of unique reads per ta's hit

ggplot(summary, aes(y=tas_hit, x=template_reads, label=sample)) +
  geom_point() + 
  geom_text(hjust=0.5, vjust=-0.5) +
  geom_smooth(method="lm", se=F)
ggsave(here("menadione_tnseq/images/reads_v_taHit.png"))
#or can use ins_density, same plot
```

No obvious correlation with number of non-duplicate template reads and number of TA's hit

## saturation curves of number of TA's hit with adding technical and biological replicates

Does including the replicates increase the number of unique TA sites hit for each experiment/biological replicate? Would pooling the biological replicates increase number of unique TA sites hit?

do separately for each condition. Read in summary data

```{r summary_plot}
library(dplyr)
library(ggplot2)
summary <- read_csv("output/sample_stats.csv", col_names=T, show_col_types = F)
summary$menadione <- as.factor(summary$menadione)
summary$experiment <- as.factor(summary$experiment)

men_plus  <- summary %>% 
  filter(menadione == 1) %>% 
  select(c(sample, experiment, ta_reads, nz_mean))
men_minus <- summary %>% 
  filter(menadione == 0) %>% 
  select(c(sample, experiment, ta_reads, nz_mean))

men_plus %>% filter(experiment==1)

ggplot(data=summary, aes(x=factor(sample, levels=c("A1", "C1", "B1", "D1","A2","C2","B2","D2")), y=ta_reads, fill=menadione)) +
  geom_bar(stat="identity") +
  xlab("dataset")

ggsave(here("menadione_tnseq/images/insertion_density.png"))

```

Do the number of unique TA sites hit increase with adding the replicates or by pooling the biological experiments?

Get all insertion data into dataframe

```{r read_wigfiles_into_dataframe}

library(purrr)
# sample name, condition, experiment, ta_site, read_count
data_path = here('menadione_tnseq/output')
#wig_files <- dir(data_path, pattern="*.wig$")
wig_files = list.files(path = data_path, pattern = "_insertions.wig$", full.names = FALSE)

df1 <- tibble(sample=wig_files) %>%
  mutate(file_contents = map(wig_files,         
           ~ read_delim(file.path(data_path, .), skip=1, delim=" ", show_col_types = F))
        )  
colnames(unnest(df1, cols=2))
# rename with sample name instead of filename (have to do this for specific samples) maybe rename in directory first
df1[[1]]<-sub("_insertions.wig", "", df1[[1]])

#remove repeat tnseq samples
men_df <- df1[1:8,]

# list of nested dfs
unnest(men_df, cols = 1)

# add columns to nested df
men_df <- men_df %>% mutate("condition" = c("men+", "men+", "men-", "men-", "men+", "men+", "men-", "men-"), "experiment" = c(rep(c("1", "2"),4)))


# all unnested
men_df1 <- unnest(men_df, cols=2)
#change column names
men_df1 <- rename(men_df1, ta_coord = variableStep, 
       reads = `chrom=Mbovis_AF2122_97`)

# get particular sample
unnest(men_df1, cols = 2) %>%
  filter(sample=="D1")

#save as R object
saveRDS(men_df1, here("menadione_tnseq/R_data/men_data.Rdata"))

```


Find set of unique sites hit for each dataset. plot sum of this set after adding more sites from replicates.

```{r sat_curves_men_positive}
library(dplyr)
library(gridExtra)
men_df1 <- readRDS(here("menadione_tnseq/R_data/men_data.Rdata"))

#for each sample, make set of non-zero ta sites
nz_sites <- men_df1 %>% filter(reads != 0)

nz_sites_plus <- nz_sites %>% filter(condition=="men+")
nz_sites_minus <- nz_sites %>% filter(condition=="men-")

A1 <- nz_sites_plus %>% filter(sample=="A1") %>% select(ta_coord) %>% pull()
A2 <- nz_sites_plus %>% filter(sample=="A2") %>% select(ta_coord) %>% pull()
C1 <- nz_sites_plus %>% filter(sample=="C1") %>% select(ta_coord) %>% pull()
C2 <- nz_sites_plus %>% filter(sample=="C2") %>% select(ta_coord) %>% pull()

#find total insertion density for each experiment
ins_den_exp1_positive <- length(union(A1,C1)) / total_ta
#0.50
ins_den_exp2_positive <- length(union(A2, C2)) / total_ta
#0.27

#combine reps for each library incrementally
exp1 <- union(A1, C1)
exp2 <- union(exp1, A2)
# all together
plus_union <- union(exp2, C2)

#number of unique ta sites with all together
length(plus_union)
#39304
#insertion density (/73536)
total_ta = 73536
length(plus_union)/total_ta
#0.53

#number of total reads with insertions in all reps/exp (union) vs number in common in all samples (intersect)
int_ta <- intersect(intersect(A1,C1), intersect(A2,C2))
length(int_ta)/ length(plus_union)
#0.25
#25% of TAs in total hit TAs are in all 4 datasets


#make df
pos_cumul_unique <- tibble(
  dataset=c("A1", "C1", "A2", "C2"),
  exp = c(1,1,2,2),
  len_tas = c(length(A1), length(C1), length(A2), length(C2)),
  cum_tas = c(length(A1), length(exp1), length(exp2), length(plus_union)),
  cum_ins_den = c(length(A1)/total_ta, length(exp1)/total_ta, length(exp2)/total_ta, length(plus_union)/total_ta)
  )


#plot for exp 1
data_p1 = pos_cumul_unique %>% 
  filter(exp==1) %>% 
  select(dataset, len_tas) %>%
  mutate(cum_tas = c(length(A1), length(union(A1, C1))) ) %>%
  mutate(ins_den = c(length(A1)/total_ta, length(C1)/total_ta) ) %>%
  mutate(cum_ins_dens = c(length(A1)/total_ta, length(union(A1, C1))/total_ta))

pl_exp1 <- ggplot(data_p1, aes(x=dataset, y=ins_den)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_ins_dens), group=1) +
  geom_point(aes(y=cum_ins_dens), group=1) +
  ylab("Insertion density") +
  xlab("experiment 1") +
  ggtitle(c("Menadione positive, experiment 1"))
  
#plot exp 2
data_p2 = pos_cumul_unique %>% filter(exp==2) %>% 
  select(dataset, len_tas) %>%
  mutate(cum_tas = c(length(A2), length(union(A2, C2))) ) %>%
  mutate(ins_den = c(length(A2)/total_ta, length(C2)/total_ta) ) %>%
  mutate(cum_ins_dens = c(length(A2)/total_ta, length(union(A2, C2))/total_ta))
pl_exp2 <- ggplot(data_p2, aes(x=dataset, y=ins_den)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_ins_dens), group=1) +
  geom_point(aes(y=cum_ins_dens), group=1) +
  ylab("insertion density") +
  xlab("experiment 2") +
  ylim(0, 0.5) +
  ggtitle(c("Menadione positive, experiment 2"))

#plot with all reps both libraries--insertion density
pos_cumul_unique <- pos_cumul_unique %>% mutate(ins_den = len_tas/total_ta)
pl_pos <- ggplot(pos_cumul_unique, 
                 aes(x=factor(dataset, levels=c("A1", "C1", "A2", "C2")),
                     y=ins_den)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_ins_den), group=1) +
  geom_point(aes(y=cum_ins_den), group=1) +
  ylab("Insertion density") +
  xlab("") +
  ggtitle("Menadione positive both libraries")

pl_pos

pl_pos_sites <- ggplot(pos_cumul_unique, 
                 aes(x=factor(dataset, levels=c("A1", "C1", "A2", "C2")),
                     y=len_tas)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_tas), group=1) +
  geom_point(aes(y=cum_tas), group=1) +
  ylab("Number of unique TA sites") +
  xlab("") +
  ggtitle("Menadione positive both libraries")

pl_pos_sites

pos_plots <- grid.arrange(pl_exp1, pl_exp2, pl_pos, pl_pos_sites)
ggsave(here("menadione_tnseq/images/cumul_insertion_pos.png"), pos_plots)

#best replicate from each experiment (A1 and A2)
best_data <- tibble(dataset=c("A1", "A2"),
  len_tas = c(length(A1), length(A2)),
  ins_den = c(length(A1)/total_ta, length(A2)/total_ta),
  cum_tas = c(length(A1), length(union(A1, A2)) ),
  cum_ins_den = cum_tas/total_ta
  )

pl_best_rep <- ggplot(best_data, 
                 aes(x=factor(dataset, levels=c("A1","A2")),
                     y=ins_den)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_ins_den), group=1) +
  geom_point(aes(y=cum_ins_den), group=1) +
  ylab("Insertion density") +
  xlab("") +
  ggtitle("Menadione positive best reps from each experiment")

pl_best_rep

```


```{r sat_curves_men_neg}
library(dplyr)
library(gridExtra)
men_df1 <- readRDS(here("menadione_tnseq/R_data/men_data.Rdata"))
#for each sample, make set of non-zero ta sites
nz_sites <- men_df1 %>% filter(reads != 0)
nz_sites_minus <- nz_sites %>% filter(condition=="men-")

#menadione minus
B1 <- nz_sites_minus %>% filter(sample=="B1") %>% select(ta_coord) %>% pull()
B2 <- nz_sites_minus %>% filter(sample=="B2") %>% select(ta_coord) %>% pull()
D1 <- nz_sites_minus %>% filter(sample=="D1") %>% select(ta_coord) %>% pull()
D2 <- nz_sites_minus %>% filter(sample=="D2") %>% select(ta_coord) %>% pull()

#combine reps for each library
exp1m <- union(B1, D1)
exp2m <- union(exp1m, B2)
# all together
plus_union_m <- union(exp2m, D2)

#number of unique ta sites with all together
length(plus_union_m)
#46597
#insertion density (/73536)
total_ta = 73536
length(plus_union_m)/total_ta
#0.63

#number of total reads with insertions in all reps/exp (union) vs number in common in all samples (intersect)
int_ta_m <- intersect(intersect(B1,D1), intersect(B2, D2))
length(int_ta_m)/ length(plus_union_m)
#0.22
#22% of TAs in total hit TAs are in all 4 datasets


#make df
neg_cumul_unique <- tibble(
  dataset=c("B1", "D1", "B2", "D2"),
  exp = c(1,1,2,2),
  len_tas = c(length(B1), length(D1), length(B2), length(D2)),
  cum_tas = c(length(B1), length(exp1m), length(exp2m), length(plus_union_m)),
  cum_ins_den = c(length(B1)/total_ta, length(exp1m)/total_ta, length(exp2m)/total_ta, length(plus_union_m)/total_ta)
  )


#plot for exp 1
data_m1 = neg_cumul_unique %>% filter(exp==1) %>% select(dataset, len_tas) %>%
  mutate(cum_tas = c(length(B1), length(union(B1, D1))) ) %>%
  mutate(ins_den = c(length(B1)/total_ta, length(D1)/total_ta) ) %>%
  mutate(cum_ins_den = c(length(B1)/total_ta, length(union(B1, D1))/total_ta))
pl_exp1m <- ggplot(data_m1, aes(x=dataset, y=ins_den)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_ins_den), group=1) +
  geom_point(aes(y=cum_ins_den), group=1) +
  ylab("Insertion density") +
  xlab("experiment 1") +
  ggtitle(c("Menadione negative, experiment 1"))
  
#plot exp 2
data_m2 = pos_cumul_unique %>% filter(exp==2) %>% 
  select(dataset, len_tas) %>%
  mutate(cum_tas = c(length(B2), length(union(B2, D2))) ) %>%
  mutate(ins_den = c(length(B2)/total_ta, length(D2)/total_ta) ) %>%
  mutate(cum_ins_den = c(length(B2)/total_ta, length(union(B2, D2))/total_ta))
pl_exp2m <- ggplot(data_m2, aes(x=dataset, y=ins_den)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_ins_den), group=1) +
  geom_point(aes(y=cum_ins_den), group=1) +
  ylab("Insertion density") +
  xlab("experiment 2") +
  ggtitle(c("Menadione negative, experiment 2"))

#plot with all reps both libraries--insertion density
neg_cumul_unique <- neg_cumul_unique %>% mutate(ins_den = len_tas/total_ta)
pl_neg <- ggplot(neg_cumul_unique, 
                 aes(x=factor(dataset, levels=c("B1", "D1", "B2", "D2")),
                     y=ins_den)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_ins_den), group=1) +
  geom_point(aes(y=cum_ins_den), group=1) +
  ylab("Insertion density") +
  xlab("") +
  ggtitle("Menadione negative both libraries")

pl_neg

pl_neg_sites <- ggplot(neg_cumul_unique, 
                 aes(x=factor(dataset, levels=c("B1", "D1", "B2", "D2")),
                     y=len_tas)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_tas), group=1) +
  geom_point(aes(y=cum_tas), group=1) +
  ylab("Number of unique TA sites") +
  xlab("") +
  ggtitle("Menadione negative both libraries")

pl_neg_sites

neg_plots <- grid.arrange(pl_exp1m, pl_exp2m, pl_neg, pl_neg_sites)
ggsave(here("menadione_tnseq/images/cumul_insertion_neg.png"), neg_plots)

#best replicate from each experiment (D1 and B2)
best_data_m <- tibble(dataset=c("D1", "B2"),
  len_tas = c(length(D1), length(B2)),
  ins_den = c(length(D1)/total_ta, length(B2)/total_ta),
  cum_tas = c(length(D1), length(union(D1, B2)) ),
  cum_ins_den = cum_tas/total_ta
  )

pl_best_rep_m <- ggplot(best_data_m, 
                 aes(x=factor(dataset, levels=c("D1","B2")),
                     y=ins_den)) +
  geom_bar(stat="identity") +
  geom_line(aes(y=cum_ins_den), group=1) +
  geom_point(aes(y=cum_ins_den), group=1) +
  ylab("Insertion density") +
  xlab("") +
  ggtitle("Menadione negative best reps from each experiment")

pl_best_rep_m

# plots grouped by experiment
grid.arrange(pl_exp1, pl_exp1m, pl_exp2, pl_exp2m)

```
This seems to show that we have a sampling bottleneck somewhere (esp in men positive) but should be random as each dataset has different set of insertion sites. Menadione neg has much better coverage/insertion density.

Definite increase in sites when adding two libraries.

May want to consider summing reads at sites between replicates rather than mean as men negative exp1 has one very sparse replicate. check this in HMM--will it take mean with zeros or just non-zero counts?

Barplot of insertion densities for each sample

```{r insertion_density plot}
library(dplyr)
men_df1 <- readRDS(here("menadione_tnseq/R_data/men_data.Rdata"))
total_ta = 73536
#non-zero ta sites
nz_sites <- men_df1 %>% filter(reads != 0)
nz_sum <- nz_sites %>% group_by(sample) %>% select(ta_coord) %>% count() %>% mutate(insertion_density = n/total_ta) %>% ungroup() %>%mutate(condition = c("men+", "men+", "men-", "men-", "men+", "men+", "men-", "men-")) 
ggplot(nz_sum, aes(x=sample, y=insertion_density, fill=condition)) +
  geom_bar(stat="identity")
ggsave(here("menadione_tnseq/images/insertion_density.png"))
```



### Look at difference in number of unique sites between conditions in each experiment

```{r exp_differences}
total_ta = 73536

#exp 1
#pos
exp1p <- union(A1,C1)
#neg
exp1m <- union(B1, D1)

#insertion densities of pooled
exp1p_dens <- length(exp1p)/total_ta
exp1m_dens <- length(exp1m)/total_ta
#difference in pooled insertion densities
exp1m_dens - exp1p_dens
#0.021

#difference in number of unique sites
diff_exp1 <- length(exp1m) - length(exp1p)
diff_exp1
#1565

# exp 2
exp2p <- union(A2, C2)
exp2m <- union(B2, D2)
diff_exp2 <- length(exp2m) - length(exp2p)
diff_exp2
#18775

#ratio of difference
diff_exp2/diff_exp1

#insertion densities of pooled
exp2p_dens <- length(exp2p)/total_ta
exp2m_dens <- length(exp2m)/total_ta
#difference in pooled insertion densities
exp2m_dens - exp2p_dens
#0.25

```

Experiment 2 has > 10X greater difference in the number of unique insertions between the conditions than experiment 1. The difference in the pooled insertion densities is 0.02 for experiment 2, but > 10X more for experiment 2 (0.26). 26% reduction seems very high for a viable menadione concentration.

Experiment 2 has more of a bottleneck in the men-positive cultures? Could this be experimental error (higher menadione concentration?) or sampling bottlenecks? Both A2 and C2 had >10M mapped reads (reads with transposon tags), indicating good amplification of the transposon junction, but very low diversity of the insertion sites (both ~20%).


## distribution of each dataset--both pooled with tech reps and without

```{r histogram_qc}
# histogram of nz-reads

men_df1 <- readRDS(here("menadione_tnseq/R_data/men_data.Rdata"))

for (i in 1:8){
  nxt_sample = men_df$sample[i]
  # get particular sample
  sample_df <- unnest(men_df1, cols = 2) %>% filter(sample==nxt_sample)
  sample_nz <- sample_df %>% filter(reads!=0) %>% select(ta_coord, reads)
  pl <- ggplot(sample_nz, aes(x=reads)) +
    geom_histogram(bins=1000) +
    xlim(1,2000) +
    ylim(0, 1000) +
    ggtitle(paste(nxt_sample, " distribution of non-zero reads"))
  file <- paste("menadione_tnseq/images/", nxt_sample, "_nz_distribution.pdf", sep="")
  ggsave(here(file), pl)
  #qqplot(y=sample_nz$reads, x=qgeom(ppoints(length(sample_nz$reads)), prob= 1/(mean(sample_nz$reads))) )
  }



a1_nz <- men_df1 %>% filter(sample=="A1" & reads!=0)
ggplot(a1_nz, aes(x=reads)) +
  geom_histogram(bins=1000) +
    xlim(1, 2000) +
    ylim(0, 1000) +
    ggtitle(paste("A1 distribution of non-zero reads"))
ggsave(here("menadione_tnseq/images/A1_histogram.png"))
max(a1_nz$reads)
min(a1_nz$reads)

# qq plot (how well does distribution resemble geometric distribution)
#qqnorm(y=a1_nz$reads)
png(here("menadione_tnseq/images/qq_plot_A1.png"))
qq <- qqplot(y=a1_nz$reads, x=qgeom(ppoints(length(a1_nz$reads)), prob= 1/(mean(a1_nz$reads))), xlab = "Data Quantiles", ylab = "Theoretical Quantiles")
dev.off()

# rank frequency
# plot read counts in sorted order
a2 <- men_df1 %>% filter(sample=="A2" & reads!=0)
a2_sorted <- sort(a2$reads, decreasing=F)
a2_ranks <- rank(a2_sorted, na.last=T)
plot(a2_sorted, a2_ranks, xlab="non-zero read counts")


#scipy.stats.probplot(truncnzreads, dist="geom", sparams=(1.0/numpy.mean(truncnzreads),))

```
What is nz_median and how does it compare to nz_mean?

```{r stats}
library(dplyr)
library(moments)
men_df1 <- readRDS(here("menadione_tnseq/R_data/men_data.Rdata"))

stat_df <- tibble(sample=men_df$sample, nz_mean=0, nz_median=0, sample_skew=0, sample_kurt=0)
for (i in 1:8){
  nxt_sample = men_df$sample[i]
  # get particular sample
  sample_df <- unnest(men_df1, cols = 2) %>% filter(sample==nxt_sample)
  sample_nz <- sample_df %>% filter(reads!=0) %>% select(ta_coord, reads)
  stat_df$nz_mean[i]     <- round(mean(sample_nz$reads),4)
  stat_df$nz_median[i]   <- median(sample_nz$reads)
  stat_df$sample_skew[i] <- round(skewness(sample_nz$reads),4)
  stat_df$sample_kurt[i] <- round(kurtosis(sample_nz$reads),4)
}
filename <- "menadione_tnseq/output/stats.tsv"
write_delim(stat_df, here(filename), delim="\t")

stat_df
```

How do the replicates and the libraries correlate? Can do this by insertions or by genes with insertions.

Plot coordinates with non-zero reads against each other? NO
Plot reads at all coordinates

```{r coord_correlation}
library(dplyr)
library(ggplot2)
library(gridExtra)
#read combined wig file
wig_data <- read_delim(here("menadione_tnseq/output/combined_samples_nonorm.wig"), delim="\t", comment="#", col_names = F)
names(wig_data) <- c("site", "A1", "C1", "A2", "C2","B1","D1","B2","D2", "gene")
wig_data_log <- wig_data %>% 
  mutate(across(A1:D2, ~ ifelse(.x != 0, log10(.x), 0), 
                .names = '{paste("log", .col, sep="")}'))
p1 <- ggplot(wig_data_log) +
  geom_point(aes(x=logA1, y=logC1))  +
  xlim(0,5) +
  ylim(0,5)
p2 <- ggplot(wig_data_log) + 
  geom_point(aes(x=logA2, y=logC2) ) +
  xlim(0,5) +
  ylim(0,5)
m1 <- ggplot(wig_data_log) + 
  geom_point(aes(x=logB1, y=logD1)) +
  xlim(0,5) +
  ylim(0,5)
m2 <- ggplot(wig_data_log) + 
  geom_point(aes(x=logB2, y=logD2)) +
  xlim(0,5) +
  ylim(0,5)

grid.arrange(p1, p2, m1, m2)


#get sites and what gene they are assigned to from one of the combined wig files

gene_insertions <- wig_data %>% select(c(2:10)) %>% group_by(gene) %>% summarise(across(everything(), sum))

#saveRDS(gene_insertions, here("menadione_tnseq/R_data/per_gene_reads.RData"))


range(gene_insertions$A1)
range(gene_insertions$C1)
nrow(gene_insertions %>% filter(A1 < 200000))
#this filters out 2 largest read counts in A1
p1g <- ggplot(gene_insertions) +
  geom_point(aes(x=log10(A1), y=log10(C1))) +
  xlim(0,6) +
  ylim(0,6)

p2g <- ggplot(gene_insertions ) +
  geom_point(aes(x=log10(A2), y=log10(C2))) +
  xlim(0,6) +
  ylim(0,6)

m1g <- ggplot(gene_insertions) +
  geom_point(aes(x=log10(B1), y=log10(D1)) )+
  xlim(0,6) +
  ylim(0,6)

m2g <- ggplot(gene_insertions) +
  geom_point(aes(x=log10(B2), y=log10(D2))) +
  xlim(0,6) +
  ylim(0,6)

cor_plot <- grid.arrange(p1g, p2g, m1g, m2g)
ggsave(here("menadione_tnseq/images/cor_reps.png"), cor_plot)
```

The level of correlation seems related to the number of reads sequenced in the dataset when using raw data. Using log transformed gives better representation. 

```{r genes_with_insertions}
library(dplyr)
gene_insertions <- readRDS(here("menadione_tnseq/R_data/per_gene_reads.RData"))

#how many genes have insertions? (genes with sum of all insertions==0)
gene_insertions <- gene_insertions %>% 
  mutate(sum = rowSums(across(where(is.numeric))))

nrow(gene_insertions %>% filter(sum!=0))
#3923 with insertions total
nrow(gene_insertions %>% filter(sum==0))
#321 without insertions total

#sort by condition
pos_gene_insertions <- gene_insertions %>% 
  select(c(gene, A1, A2, C1, C2)) %>% 
  mutate(sum = rowSums(across(where(is.numeric))))
nrow(pos_gene_insertions %>% filter(sum!=0))
#3793
nrow(pos_gene_insertions %>% filter(sum==0))
#451

neg_gene_insertions <- gene_insertions %>% 
  select(c(gene, B1, B2, D1, D2)) %>%
  mutate(sum = rowSums(across(where(is.numeric))))
nrow(neg_gene_insertions %>% filter(sum!=0))
#3867
nrow(neg_gene_insertions %>% filter(sum==0))
#377

total_genes = nrow(gene_insertions)

gene.df <- tibble(condition = c("negative", "positive"),
                  inserted_genes = c(
                    #nrow(gene_insertions %>% filter(sum!=0)),
                    nrow(neg_gene_insertions %>% filter(sum!=0)),
                    nrow(pos_gene_insertions %>% filter(sum!=0))
                    ) 
                  )

ggplot(gene.df) +
  geom_bar(stat="identity", aes(x=condition, y=inserted_genes)) +
  ylim(c(0,5000)) +
  ggtitle("Comparison of number of genes with insertions")

gene_none.df <- tibble(condition = c("both", "negative", "positive"),
                  inserted_genes = c(
                    nrow(gene_insertions %>% filter(sum==0)),
                    nrow(neg_gene_insertions %>% filter(sum==0)),
                    nrow(pos_gene_insertions %>% filter(sum==0))
                    ) 
                  )
ggplot(gene_none.df) +
  geom_bar(stat="identity", aes(x=condition, y=inserted_genes)) +
  
  ggtitle("Comparison of number of genes with no insertions")

# % of genes with insertions in each
total_genes = nrow(gene_insertions)
data <- tibble(condition = c("negative", "positive"), 
               inserted_genes = c(nrow(neg_gene_insertions %>% filter(sum!=0)), 
                                nrow(pos_gene_insertions %>% filter(sum!=0))),
               all_genes = c(total_genes, total_genes)
               )
data <- data %>% mutate(percent_inserted = round(inserted_genes/total_genes,2)*100)

ggplot(data, aes(x=condition, y=all_genes, fill=inserted_genes)) +
  geom_bar(stat="identity")

```


I am more interested in correlation between TA sites hit. Change values to 0 or 1 if there is non-zero

```{r ta_site_binary_correlation}

library(dplyr)
library(ggplot2)
library(gridExtra)
#read combined wig file
wig_data <- read_delim(here("menadione_tnseq/output/combined_samples_nonorm.wig"), delim="\t", comment="#", col_names = F, show_col_types = FALSE)
names(wig_data) <- c("site", "A1", "C1", "A2", "C2","B1","D1","B2","D2", "gene")
wig_binary <- wig_data %>% mutate(across(A1:D2, ~ ifelse(.x != 0, 1, 0)) )

#can't plot, just looks like four points ((0,1), (0,0), (1,0), (1,1))


```



From Chouderey et al, 2021 (September/October 2021 Volume 6 Issue 5 e00876-21 msystems.asm.org 22
Nucleotide Biases Affecting Himar Transposon Insertion):

>The correlation of insertion counts at TA sites between TnSeq data sets was calculated using a Pearson correlation coefficient. As mentioned previously, the log of insertion counts was used, since the Pearson correlation coefficient assumes that the input data are normally distributed. The two-tailed t test for the means for two independent samples was used to measure whether the expected value differs significantly between samples. Since we do not assume population variance between the two data sets is equal, the Welch t test is performed

This is log10 as it better fits Gaussian distribution

this can be done between libraries, as well

```{r correlation between datasets}
library(dplyr)

wig_data <- read_delim(here("menadione_tnseq/output/combined_samples_nonorm.wig"), delim="\t", comment="#", col_names = F, show_col_types = FALSE)
names(wig_data) <- c("site", "A1", "C1", "A2", "C2","B1","D1","B2","D2", "gene")

#using log10 of insertion counts 
#mutate(across(starts_with("test_"), ~ .x == get(sub("test_", "", cur_column())), 
     #           .names = '{gsub("test_", "answer_", .col)}'))

#can't use -Inf so changed to 0 if no reads (log10(1)==0 so looks like a lot more insertions with 1 count, but that's ok for distribution?)
wig_data_log <- wig_data %>% 
  mutate(across(A1:D2, ~ ifelse(.x != 0, log10(.x), 0), 
                .names = '{paste("log", .col, sep="")}'))

p <-  cor.test(wig_data_log$logA1, wig_data_log$logC1, method="spearman")
p$estimate[[1]]
#0.68
cor.test(wig_data_log$logA2, wig_data_log$logC2, method="spearman")
#0.89
cor.test(wig_data_log$logB1, wig_data_log$logD1, method="spearman")
#0.52
cor.test(wig_data_log$logB2, wig_data_log$logD2, method="spearman")
#0.76

#test between libraries (A1 v C2, A1 v A2, A2 v C1)
cor.test(wig_data_log$logA1, wig_data_log$logC2, method = "spearman")
#0.36, 0.35, 0.33
#B1vB2, B1 v D2, B2 vD1
cor.test(wig_data_log$logB2, wig_data_log$logD1)
#0.32, 0.30, 0.52(best between B2 and D1 which are best saturated)

#two tailed t-test to see if expected value of insertions differs between replicates (did both men + and -)
t.test(wig_data_log$logA2, wig_data_log$logC2, var.equal=F)
#all of them support alt hypothesis (that true difference in means is not equal to 0) and therefore the populations from which the samples were taken have different means, i.e. they don't really look like they are from same overall population

#did same between libraries (one rep only)
t.test(wig_data_log$logB2, wig_data_log$logD1, var.equal=F)
#all libraries show difference in means is more than expected if these are coming from same population which is to be expected for different libraries


```
Insertion counts at the same site in different libraries is expected to be uncorrelated because of so many sources of variation--indicates not influenced by insertion bias.

I would expect replicates to be better correlated (and they are) because replicates do not have as much variation. If they are not correlated, seems to indicate bottleneck in downstream processing: sampling the gDNA, etc.

```{r barcharts_unique insertions}
#look at sets of TA coord between reps

#for each sample, make set of non-zero ta sites

A1 <- wig_data %>% filter(A1 != 0) %>% select(site) %>% pull()
A2 <- wig_data %>% filter(A2 !=0) %>% select(site) %>% pull()
C1 <- wig_data %>% filter(C1 !=0) %>% select(site) %>% pull()
C2 <- wig_data %>% filter(C2 !=0) %>% select(site) %>% pull()
B1 <- wig_data %>% filter(B1 != 0) %>% select(site) %>% pull()
B2 <- wig_data %>% filter(B2 != 0) %>% select(site) %>% pull()
D1 <- wig_data %>% filter(D1 != 0) %>% select(site) %>% pull()
D2 <- wig_data %>% filter(D2 != 0) %>% select(site) %>% pull()

un_ta <- tibble(A1_sites = length(A1), C1_sites=length(C1), both_exp1 = length(intersect(A1,C1)), A2_sites = length(A2), C2_sites=length(C2), both_exp2=length(intersect(A2, C2)), in_all=length(intersect(intersect(A1, A2), intersect(C1,C2))))
d_unta <- pivot_longer(un_ta, cols= A1_sites:in_all, names_to="sample", values_to="unique_sites") %>% mutate(exp = c(rep("exp1", 3), rep("exp2", 3), "all" ) )
d_unta$sample <- as.factor(d_unta$sample)

ggplot(d_unta, aes(fill=exp)) +
  geom_bar(stat="identity", aes(x=factor(sample, levels=c("A1_sites", "C1_sites", "both_exp1", "A2_sites", "C2_sites", "both_exp2", "in_all")), y=unique_sites)) +
  xlab("Samples") +
  ggtitle("Menadione Positive unique TA sites") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave(here("menadione_tnseq/images/unique_insertions_pos.png"))

#number of total reads with insertions in all reps/exp (union) vs number in common in all samples (intersect)
int_ta <- intersect(intersect(A1,C1), intersect(A2,C2))
union_ta <- union(union(A1,C1), union(A2,C2))
length(int_ta)/ length(union(union(A1,C1), union(A2,C2)))
#0.25
#25% of TAs in total hit TAs are in all 4 datasets

#plot with ratios
data <- d_unta %>% mutate(ratio = round(unique_sites/length(union_ta),2))

p1 <- ggplot(data, aes(fill=exp)) +
  geom_bar(stat="identity", aes(x=factor(sample, levels=c("A1_sites", "C1_sites", "both_exp1", "A2_sites", "C2_sites", "both_exp2", "in_all")), y=ratio)) +
  xlab("Samples") +
  ylab("fraction of total unique sites") +
  ggtitle("Menadione Positive unique TA sites") +
  theme(plot.title = element_text(hjust = 0.5))


#for control

un_ta_m <- tibble(B1_sites = length(B1), D1_sites=length(D1), both_exp1 = length(intersect(B1,D1)), B2_sites = length(B2), D2_sites=length(D2), both_exp2=length(intersect(B2, D2)), in_all=length(intersect(intersect(B1, B2), intersect(D1,D2))))
d_unta_m <- pivot_longer(un_ta_m, cols= B1_sites:in_all, names_to="sample", values_to="unique_sites") %>% mutate(exp = c(rep("exp1", 3), rep("exp2", 3), "all" ) )
d_unta_m$sample <- as.factor(d_unta_m$sample)

ggplot(d_unta_m, aes(fill=exp)) +
  geom_bar(stat="identity", aes(x=factor(sample, levels=c("B1_sites", "D1_sites", "both_exp1", "B2_sites", "D2_sites", "both_exp2", "in_all")), y=unique_sites)) +
  xlab("Samples") +
  ggtitle("Menadione Negative unique TA sites") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave(here("menadione_tnseq/images/unique_insertions_neg.png"))

#plot with ratios

#number of total reads with insertions in all reps/exp (union) vs number in common in all samples (intersect)
int_ta_m <- intersect(intersect(B1,D1), intersect(B2,D2))
union_ta_m <- union(union(B1,D1), union(B2,D2))
length(int_ta_m)/ length(union(union(B1,D1), union(B2,D2)))
#0.21
#21% of TAs in total hit TAs are in all 4 datasets

data <- d_unta_m %>% mutate(ratio = round(unique_sites/length(union_ta),2))

p1 <- ggplot(data, aes(fill=exp)) +
  geom_bar(stat="identity", aes(x=factor(sample, levels=c("B1_sites", "D1_sites", "both_exp1", "B2_sites", "D2_sites", "both_exp2", "in_all")), y=ratio)) +
  xlab("Samples") +
  ylab("fraction of total unique sites") +
  ggtitle("Menadione Negative unique TA sites") +
  theme(plot.title = element_text(hjust = 0.5))


```

Use corplot to make correlation plots like used in TRANSIT

Use the mean insertion counts (averaged at the gene-level) among samples

```{r corplots}
library(corrplot)

#this is messy--overlapping TA sites for more than one gene will have to be dealt with
# wig_data <- read_delim(here("menadione_tnseq/output/combined_samples_nonorm.wig"), delim="\t", comment="#", col_names = F, show_col_types = FALSE)
# names(wig_data) <- c("site", "A1", "C1", "A2", "C2","B1","D1","B2","D2", "gene")
# 
# #make mean of reads per gene
# mean_genes <- wig_data %>% select(c(2:10)) %>% group_by(gene) %>% summarise(across(everything(), mean))

#use file generated by transit export mean_genes function (same issue)

gene_means <- read_tsv(here("menadione_tnseq/output/combined_gene_means.txt"), col_names = F, comment="#", show_col_types = F)
names(gene_means) <- c("ORF", "name", "sites", "A1", "C1", "A2", "C2", "B1", "D1", "B2", "D2")	
gm_matrix <- as.matrix(gene_means %>% select(A1:D2))
#gm_matrix <- as.matrix(gene_means)
png(here("menadione_tnseq/images/corrplot_samples.png"))
corrplot(cor(gm_matrix, method="pearson"))
dev.off()


```
Differences in replicates most likely due to sampling differences when scraping plates and homogenizing cells to extract DNA, taking small portion of total gDNA for sequencing, shearing, technical errors making sequencing library, PCR jackpots (eliminated after the fact), etc. 

Insertion counts at the same site in different libraries is expected to be uncorrelated because of so many sources of variation—indicates transposon insertion not influenced by insertion bias.

